---
title: "deliverable2"
author: "Jerusalem Moore"
date: "11/27/2019"
output: html_document
---

The predictions I would like to model from the initial data would be exactly how much chroni absenteeism and more generally student performance is effected by different factors. I feel these predictions matter because they can provide an outlook on whether or not these factors play any positive or negative roles in student class participation and success.

Note:You need tidyverse, readr, and readxl installed for the following code to run


```{r}
library("readr")
library("tidyverse")
library("readxl")
suspension_data <- read_delim("suspension_data.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
chronic_absenteeism <- read_delim("C:/Users/jerus/Downloads/chrabs1718 (1).txt", "\t", escape_double = FALSE, trim_ws = TRUE)
free_reduced_meals <- read_excel("free_reduced_meals.xlsx", sheet = "FRPM School-Level Data ", skip = 1)


```

I'm going to minimize the three datasets into smaller tables, starting with an absenteeism table from the chronic absenteeism dataset

Notes for absenteeism table...

I'm going to include enrollment numbers, charter school indicators, and absenteeism rates/counts to monitor student performance and behavior

non obvious variables:
scope:indicator for whether entry is an observation by school(S), state(T), district(D), or county(C) level
charter_school:indicator for whether entry is a charter school

```{r}
absenteeism_table <- tibble(year = chronic_absenteeism$AcademicYear, scope = chronic_absenteeism$AggregateLevel, county_code = chronic_absenteeism$CountyCode, county_name = chronic_absenteeism$CountyName, charter_school = chronic_absenteeism$CharterYN, total_enrollment = chronic_absenteeism$CumulativeEnrollment, absenteeism_count = chronic_absenteeism$ChronicAbsenteeismCount, absenteeism_rate = chronic_absenteeism$ChronicAbsenteeismRate)
```

Next I'm going to make a suspension table from the suspension dataset in order to supplement abesenteeism which I believe will add some depth to student performance and behavior

```{r}
suspension_table <- tibble(year = suspension_data$AcademicYear, scope = suspension_data$AggregateLevel, county_code = suspension_data$CountyCode, county_name = suspension_data$CountyName, charter_school = suspension_data$CharterYN, total_enrollment = suspension_data$`Cumulative Enrollment`, total_suspensions = suspension_data$`Total Suspensions`, suspension_rate = suspension_data$`Suspension Rate (Total)`)

```

Lastly I'm going to create a table from the free_reduced_meals dataset in order to see if the percent of students eligible for free food and educational option type have any relation to school performance and behavior
```{r}
names(free_reduced_meals)[10] <- "educational_option_type"
names(free_reduced_meals)[20] <- "percent_free_food_eligible"
free_meals_table <- tibble(year = free_reduced_meals$`Academic Year`, county_code = free_reduced_meals$`County Code`, county_name = free_reduced_meals$`County Name`, education_type = free_reduced_meals$educational_option_type, percent_free_food_eligible = free_reduced_meals$percent_free_food_eligible)
```